---
title: "Visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(waveformtools)
library(tidyverse)
library(mgcv)
library(hms)
library(shiny)
library(shinythemes)
library(shinyTime)
library(data.table)
library(readr)
library(dygraphs)
library(dplyr)
library(tseries)
library(mgcViz)
```

## Load dataset

```{r}
filepath <- "../Datasets/thoracotomy_waveforms/pt18.RDS"
full_dataload <- readRDS(filepath, refhook = NULL)
```

```{r}
add_time <- function(data, ms_between_obs = 8, time_origin = "2021-09-03") {
  #' adds a column containing time with increments of ms_between_obs
  #' data:             a dataframe like structure
  #' ms_between_obs:   the amount of milliseconds between each cycle
  df <- tibble(data)
  df$time_ms <- seq(0, nrow(df)*ms_between_obs-ms_between_obs, ms_between_obs)
  df$time_s <- df$time_ms / 1000
  return(df)
}

create_patient <- function(filepath) {
  #' creates a dataframe containing cvp and time in ms and s for open chest
  #' filepath:        the path of the patient file
  dataload <- readRDS(filepath, refhook = NULL)
  patient <- data.frame(open_chest_cvp = dataload$open_chest$CVP) # extract cvp numbers from data
  patient <- add_time(patient)
  return(patient)
}
```

```{r}
pt18 <- create_patient(filepath)
qrs_ms <- full_dataload$open_chest$QRSmarker_ms

# The following code is to give an index to each measurement for which position it has in its heartbeat 
# cycle. Also removes rows containing NAs giving only rows where we know start and end of cycle
pt18 <- na.omit(gen_annotation_index(pt18, qrs_ms, 'time_ms', 'qrs')) 
pt18 <- na.omit(gen_annotation_index(pt18, seq(0, nrow(pt3)*8, 5000), 'time_ms', 'insp'))
```

### Inspect data

```{r inspecting cvp over time}
dygraph(data = pt18[,c("time_s","open_chest_cvp")])
```


```{r slicing data}
pt18_sliced <- pt18[pt18$time_s > 200 & pt18$time_s < 300,]
```

### Create gam object

```{r}
cvp_gam <- bam(
        open_chest_cvp ~ s(qrs_rel_index, bs = 'cc', k = 50) +
            s(insp_rel_index, bs = 'cc', k = 30) +
            ti(
                qrs_rel_index,
                insp_rel_index,
                bs = c('cc', 'cc'),
                k = c(20, 10)
            ) +
            s(time_s),
        method = 'REML',
        data = pt18_sliced,
        rho = 0.9,
        nthreads = 16 # Number of (virtual) cores
    )
```

### Create visualization object

```{r}
b <- getViz(cvp_gam)
```


### Conditional Probability for individual effects

Plot of position in qrs-cycle's smooth effect on cvp. Added is a layer of the conditional probabilty of the residuals given the x-values.

```{r}
plot(b, select = 1) + l_dens(type = "cond") + l_fitLine() + l_ciLine()
```

From this it can be seen that the estimate of the effect of the position in qrs-cycle often has a high precision in terms of conditional probabilty. 

Plot of position in insp-cycle's smooth effect on cvp. Added is a layer of the conditional probabilty of the residuals given the x-values.

```{r}
plot(b, select = 2) + l_dens(type = "cond") + l_fitLine() + l_ciLine()
```

Here, we can see that the estimate of the effect of the position in the inspiration-cycle is often close to zero with a high probability. This means that for this patient, when the chest is opened, the inspiration-cycle does not have a significant effect on the cvp.

### Plot of interactions-term


```{r}
plotRGL(sm(b, 3), fix = c("z" = 0), residuals = TRUE)
```

### Gam check

```{r}
gam.check(cvp_gam)
```

Consider doing a trim based on qq-plot and histogram.

```{r}
qq.gam(cvp_gam)
```

```{r}
cvp_gam_aftercheck <- bam(
        open_chest_cvp ~ s(qrs_rel_index, bs = 'cc', k = 50) +
            s(insp_rel_index, bs = 'cc', k = 30) +
            ti(
                qrs_rel_index,
                insp_rel_index,
                bs = c('cc', 'cc'),
                k = c(30, 20)
            ) +
            s(time_s),
        method = 'REML',
        data = pt18_sliced,
        rho = 0.9,
        nthreads = 16 # Number of (virtual) cores
    )
```

```{r}
gratia::draw(cvp_gam_aftercheck)
```

```{r}
c <- getViz(cvp_gam_aftercheck)
plot(c, allTerms = T)
```



```{r}
plotRGL(sm(c, 3), fix = c("z" = 0), residuals = TRUE)
```

